Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-09-30 paper.Rnw
updated text to match results in bullet example

In response, \citet{hare:2017} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas, \citet{hare:2017} obtain signatures of striations from two bullet lands (\autoref{fig:figure-09}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest (available in the \emph{bulletxtrctr} R package) to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). See the supporting information for \kge{definitions of the features \sout{additional information on the data}}. The model was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which included \kge{10,384 land-to-land comparisons \citet{hare:2017}}.

@@

We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) predictions (\autoref{fig:figure-10}). \kgc{May want to change the next sentence a bit now:} The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.

@@

# Apply LIME to two cases with the permutations returned
bullet_lime_explain_perms <- apply_lime(
  train = bullet_train %>% select(all_of(bullet_features)),
  test = bullet_test %>%
    filter(case %in% bullet_poi_match) %>%
    bind_rows(bullet_test %>%
                filter(case %in% bullet_poi_nonmatch)) %>%
    select(all_of(bullet_features)),
  model = bullet_rf,
  label = as.character(TRUE),
  n_features = 3,
  sim_method = c(
    'quantile_bins',
    'equal_bins',
    'kernel_density',
    'normal_approx'
  ),
  nbins = 3,
  feature_select = "auto",
  dist_fun = "gower",
  kernel_width = NULL,
  gower_pow = 0.5,
  return_perms = TRUE,
  all_fs = FALSE,
  seed = 20190914
)

@@

# Determine the application and case number of the poi
# in the no_perms data
poi_cases_no_perms <- bullet_explain_noperms %>%
  filter(case %in% c(bullet_poi_match, bullet_poi_nonmatch),
         nbins %in% c(3, NA),
         gower_pow == 0.5) %>%
  select(case, implementation) %>%
  unique()

# Join the no perms data (with poi removed) with the
# perms data (perms removed)
bullet_explain <- bullet_explain_noperms %>%
  anti_join(poi_cases_no_perms,
            by = c("implementation", "case")) %>%
  bind_rows(
    bullet_explain_perms %>%
      mutate(
        implementation = fct_recode(
          implementation,
          "2" = "1",
          "7" = "2",
          "31" = "3",
          "32" = "4"
        )
      ) %>%
      select(
        -perms_raw,
        -perms_numerified,-perms_pred_simple,
        -perms_pred_complex,-weights
      )
  )

@@

\kgc{Will need to edit this paragraph:} First, the density simulation methods produce the same explanations for almost all cases and LIME tuning parameter values suggesting the LIME explanations are global and not local. Second, within a bin based simulation method, the features selected by LIME for an observation often vary by the number of bins but do not appear to vary by the Gower power. With the equal bins, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations produced by the bin based simulation methods for the matches and non-matches. This suggests that the features the random forest uses to classify a match or non-match are different.

@@

@
\caption{\kgc{Need to edit this:} Assessment metric plot for the applications of LIME on the bullet comparison data test set. The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). \kgc{Need to edit the rest of this paragraph:} All density simulation methods performed well based on the three metrics. This is an interesting result, since \autoref{fig:figure-11} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend. The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME. Perhaps we can say that the kernel density method or 3 equal bins  are the best performing methods, but unfortunately, \autoref{fig:figure-12} leaves us without a clear indication of which set of explanations to use to explain the random forest.

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 3 equal bins and 3 quantile bins are included to depict tuning parameter values with good (3 equal bins) and mediocre (3 quantile bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

\kgc{Need to edit this pargraph:} First, focus on the explanation for case NM with 3 quantile bins. The explanation plot from \emph{lime} shows that an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls in and supports the random forest prediction of a non-match.

\kgc{Maybe edit this pargraph:} The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the 3 quantile bin explanations are not faithful to the random forest model predictions. The 3 equal bin explanations are more faithful to the random forest, but the regions captured by the bins still do not align with the regions with similar probabilities produced by the random forest.

@@

@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

@@

Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, a lack of understanding of the process used to create the explanation could lead to an incorrect interpretation of the explanation. For example, the visualization of a LIME explanation available from the \emph{lime} R package \citep{pedersen:2020} (shown in Figures \ref{fig:figure-04} and \ref{fig:figure-13}) is a major simplification of the explainer model, which could lead to under-interpreted or misinterpreted LIME explanation. Supplementing \citet{pedersen:2020}'s compact visualization of the explanation with an explanation scatterplot that shows a more detailed visualization of the explainer model (such as Figures~\ref{fig:figure-04} and \ref{fig:figure-13}) promotes a more complete understanding of the explanation.

\kgc{Need to make some edits in this paragraph:} Even with an explainer model that is interpreted correctly, the interpretation is worthless if the explainer model is not \textbf{faithful} to the complex model. This claim can be assessed using the diagnostic plots suggested in this paper. Many of the visualizations in this paper highlight problems with the faithfulness of the explainer models. The explanation scatterplots allow for a comparison of the explainer model to the complex model. The examples in this paper show cases where the explainer model bins do not accurately capture the regions with similar random forest probabilities that contain the prediction of interest and oversimplify the model (Figures \ref{fig:figure-04} and \ref{fig:figure-13}). Using fewer bins would clearly not help improve the faithfulness of the explainer model in these examples, and while an increase in the number of bins would lead to a finer resolution of the random forest classification boundaries, interpretability of the explainer model would quickly be lost. Perhaps this could be improved by allowing the bin creation to account for the relationships between the features and response variable or a different number of bins for each feature.

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:figure-07} and \ref{fig:figure-12}) both produced conflicting results. The density based simulation methods result in the best or close to the best performance even though the feature heatmaps (Figures \ref{fig:figure-06} and \ref{fig:figure-11}) show that the density simulation methods produce global explanations with minimal variation in features selected by LIME. For the bin based simulation methods, the two metrics often do not agree or contradict one another, which makes it difficult to decide on a recommendation of a set of tuning parameter values that produce the explanations with the most faithful explainer model.

@@

\kgc{Probably need to make some edits here:} The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plots. As stated, the feature heatmap reveals that the density simulation methods in the examples of this paper results in global explanations where the same features are repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citep{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories. Furthermore, while the metric comparison plot in the bullet example does not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is the best. This suggests that a less local explanation provides a better explanation of the performance of the random forest.

>>>>

In response, \citet{hare:2017} proposed an automated machine learning method for bullet matching to complement a visual inspection by firearm examiners. Based on high-resolution topological scans of land engraved areas, \citet{hare:2017} obtain signatures of striations from two bullet lands (\autoref{fig:figure-09}). Nine features quantifying the similarity of signatures, such as the cross-correlation function, the distance between signatures, and the number of matching striae, are extracted and used to train a random forest \kge{\sout{(available in the \emph{bulletxtrctr} R package)}} to determine the probability of a comparison resulting from the same source (matching signatures) or from different sources (non-matching signatures). The model \kge{in \citet{hare:2017}} was trained on a set of scans of bullets from the James Hamby Consecutively Rifled Ruger Barrel Study \citep{hamby:2009}, which included \kge{10,384 land-to-land comparisons \citet{hare:2017}}. \kgc{Moved this sentence to the end of the paragraph:} See the supporting information for additional information on the \kge{signature similarity features\sout{data}}.

@@

We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) \kge{\sout{predictions}} (\autoref{fig:figure-10}). \kge{\sout{The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.}} \kge{The training data observations show a clear difference in feature values based on whether the corresponding random forest probability is close to 1 or 0. For example, observations with a random forest probability close to 1 tend to have high values of rough correlation, cross correlation function, and matches and low values of mismatches, distance, and non-consecutivley matching striae. The observations in the test data that are classified incorrectly by the random forest tend to have feature values similar to observations in the training data with similar random forest scores. For example, the observations in the test set known to be a match but have a random forest score below 0.5 have relatively low values of rough correlation, cross correlation function, and matches and relatively large values of mismatches, distance, and non-consecutivley matching striae compared to observations with a random forest probability close to 1.}

@@

# Apply LIME to two cases with the permutations returned
bullet_lime_explain_perms <- apply_lime(
  train = bullet_train %>% select(all_of(bullet_features)),
  test = bullet_test %>%
    filter(case %in% bullet_poi_match) %>%
    bind_rows(bullet_test %>%
                filter(case %in% bullet_poi_nonmatch)) %>%
    select(all_of(bullet_features)),
  model = bullet_rf,
  label = as.character(TRUE),
  n_features = 3,
  sim_method = c(
    'quantile_bins',
    'equal_bins',
    'kernel_density',
    'normal_approx'
  ),
  nbins = 4,
  feature_select = "auto",
  dist_fun = "gower",
  kernel_width = NULL,
  gower_pow = 0.5,
  return_perms = TRUE,
  all_fs = FALSE,
  seed = 20190914
)

@@

# Determine the application and case number of the poi
# in the no_perms data
poi_cases_no_perms <- bullet_explain_noperms %>%
  filter(case %in% c(bullet_poi_match, bullet_poi_nonmatch),
         nbins %in% c(4, NA),
         gower_pow == 0.5) %>%
  select(case, implementation) %>%
  unique()

# Join the no perms data (with poi removed) with the
# perms data (perms removed)
bullet_explain <- bullet_explain_noperms %>%
  anti_join(poi_cases_no_perms,
            by = c("implementation", "case")) %>%
  bind_rows(
    bullet_explain_perms %>%
      mutate(implementation = fct_recode(implementation,
          "3" = "1",
          "8" = "2",
          "31" = "3",
          "32" = "4"
        )
      ) %>%
      select(
        -perms_raw,-perms_numerified,
        -perms_pred_simple,-perms_pred_complex,
        -weights
      )
  )

@@

First, \kge{2 quantile bin \sout{the density}} simulation method \kge{applications} produce the same explanations for almost all cases and LIME tuning parameter values suggesting the LIME explanations are global and not local. Second, within a \kge{simulation method \sout{bin based simulation method}}, the features selected by LIME for an observation \kge{\sout{often vary by the number of bins but}} do not appear to vary by the Gower power. \kge{However, the LIME explanation for an observation often varies across simuluation methods.} With the equal bins, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations produced by the bin based simulation methods for the matches and non-matches. This suggests that the features the random forest uses to classify a match or non-match are different.

@@

@
\caption{Assessment metric plot for the applications of LIME on the bullet comparison data test set. \kge{\sout{The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.} There are discrepancies in performance across metrics for a set of tuning parameter values, but based on the overall performance of all three metrics, 4 equal bins with a Gower power of 0.5 appears to perform the best.}}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). \kge{\sout{All density simulation methods performed well based on the three metrics. This is an interesting result, since Figure \ref{fig:figure-11} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend.}} \kge{The performance of a set of tuning parameter values often do not agree across metrics. For example, both density methods perform well according to average fidelity but poorly according to MSEE, and all quantile bin methods perform well accroding to average $R^2$ but poorly based on average fidelity and MSEE. However, there is consistency in results across different Gower powers.} The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME. \kge{\sout{Perhaps we can say that the kernel density method or 3 equal bins are the best performing methods, but unfortunately, Figure \ref{fig:figure-12} leaves us without a clear indication of which set of explanations to use to explain the random forest.}} \kge{It is not obvious which set of of tuning parameters are best, but perhaps, we can say that based on all three metrics, the 4 equal bins with a Gower power of 0.5 performs bettter than the other tuning parameter values considered.}

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 4 equal bins and 4 quantile bins are included to depict tuning parameter values with good (4 equal bins) and mediocre (4 quantile bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

First, focus on the explanation for case NM with 4 quantile bins. \kge{The explanation plot from \emph{lime} shows that observed values with rough correlation (rough\_cor) greater than 0.05 and mismatches value less than or equal to 7.17 support a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with rough correlation greater then 0.05 and mismatches less than 7.17 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of cross correlation function (ccf) greater than 0.28 and less than or equal to 0.35 supports a non-match, and the scatter plots shows that the simulated values within this region have mostly random forest predictions less than 0.5.  However, the random forest assigns a probability of 0.19 to case NM, which is not explained by the LIME explanation. Instead, consider that a better explanation based on the scatterplots would be that the random forest assigned a prediction probably close to 0 since the observation has a rough correlation less than 0.5, mismatches greater than 4, and cross correlation function 0.6.} \kge{\sout{Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls in and supports the random forest prediction of a non-match.}}

The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the \kge{\sout{3} 4} quantile bin explanations are not faithful to the random forest model predictions. The \kge{4 \sout{3}} equal bin explanations are more faithful to the random forest, but the regions captured by the bins do not align \kge{as well as they could} with the regions \kge{containing \sout{with}} similar probabilities produced by the random forest.

@@

@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: \kge{4 \sout{3}} quantile bins (first column) and \kge{4 \sout{3}} equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

@@

Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, a lack of understanding of the process used to create the explanation could lead to an incorrect interpretation of the explanation. For example, the visualization of a LIME explanation available from the \emph{lime} R package \citep{pedersen:2020} (shown in Figures \ref{fig:figure-04} and \ref{fig:figure-13}) is a major simplification of the explainer model, which could lead to under-interpreted or misinterpreted LIME explanation. Supplementing \citet{pedersen:2020}'s compact visualization of the explanation with an explanation scatterplot that shows a more detailed visualization of the explainer model (such as Figures~\ref{fig:figure-04} and \ref{fig:figure-13}) promotes a more complete understanding of the explanation.
Even with an explainer model that is interpreted correctly, the interpretation is worthless if the explainer model is not \textbf{faithful} to the complex model. This claim can be assessed using the diagnostic plots suggested in this paper. Many of the visualizations in this paper highlight problems with the faithfulness of the explainer models. The explanation scatterplots allow for a comparison of the explainer model to the complex model. The examples in this paper show cases where the explainer model bins do not accurately capture the regions with similar random forest probabilities that contain the prediction of interest and oversimplify the model (Figures \ref{fig:figure-04} and \ref{fig:figure-13}). Using fewer bins would clearly not help improve the faithfulness of the explainer model in these examples, and while an increase in the number of bins would lead to a finer resolution of the random forest classification boundaries, interpretability of the explainer model would quickly be lost. Perhaps this could be improved by allowing the bin creation to account for the relationships between the features and response variable or a different number of bins for each feature.

In addition to assessing faithfulness by visually comparing the complex and explainer model, we propose a visual comparison of two faithfulness metrics (MSEE and average fidelity). The examples of faithfulness metric comparisons in this paper (Figures \ref{fig:figure-07} and \ref{fig:figure-12}) both produced conflicting results. \kge{\sout{The density based simulation methods result in the best or close to the best performance even though the feature heatmaps (Figures \ref{fig:figure-06} and \ref{fig:figure-11}) show that the density simulation methods produce global explanations with minimal variation in features selected by LIME. For the bin based simulation methods, the two metrics often do not agree or contradict one another, which} This} makes it difficult to decide on a recommendation of a set of tuning parameter values that produce the explanations with the most faithful explainer model.

@@

The final claim, \textbf{localness}, is addressed by the feature heatmap and metric comparison plots. As stated, the feature heatmap reveals that the density simulation methods in the \kge{\data \ example \sout{examples of this paper}} results in global explanations where the same features are repeatedly chosen across all (or almost all) observations in the set of explanations. This finding agrees with that of \citet{laugel:2018} who found LIME produced global explanations with the normal approximation simulation method. For the bin based simulation methods in the bullet data example, the feature heatmap showed that the features chosen for the explanations varied between the two classification categories (match versus non-match). This is an interesting finding that suggests that different features can play a role in the predictions of observations in different response categories. Furthermore, while the metric comparison plot in the bullet example does not provide agreement between metrics on a best bin based method, all metrics agree that a Gower power of 0.5 for computing the model weights associated with distance of a simulated data point from the prediction of interest is the best. This suggests that a less local explanation provides a better explanation of the performance of the random forest.

### 2020-09-30 paper.tex
updated text to match results in bullet example

\end{knitrout}
\caption{\kgc{Need to edit this:} Assessment metric plot for the applications of LIME on the bullet comparison data test set. The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). \kgc{Need to edit the rest of this paragraph:} All density simulation methods performed well based on the three metrics. This is an interesting result, since \autoref{fig:figure-11} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend. The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME. Perhaps we can say that the kernel density method or 3 equal bins  are the best performing methods, but unfortunately, \autoref{fig:figure-12} leaves us without a clear indication of which set of explanations to use to explain the random forest.

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 3 equal bins and 3 quantile bins are included to depict tuning parameter values with good (3 equal bins) and mediocre (3 quantile bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

\kgc{Need to edit this pargraph:} First, focus on the explanation for case NM with 3 quantile bins. The explanation plot from \emph{lime} shows that an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls in and supports the random forest prediction of a non-match.

\kgc{Maybe edit this pargraph:} The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the 3 quantile bin explanations are not faithful to the random forest model predictions. The 3 equal bin explanations are more faithful to the random forest, but the regions captured by the bins still do not align with the regions with similar probabilities produced by the random forest.

@@

\end{knitrout}
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

>>>>

\end{knitrout}
\caption{Assessment metric plot for the applications of LIME on the bullet comparison data test set. \kge{\sout{The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.} There are discrepancies in performance across metrics for a set of tuning parameter values, but based on the overall performance of all three metrics, 4 equal bins with a Gower power of 0.5 appears to perform the best.}}
\label{fig:figure-12}
\end{figure*}

To try to identify a set of LIME tuning parameter values with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:figure-12}). \kge{\sout{All density simulation methods performed well based on the three metrics. This is an interesting result, since Figure \ref{fig:figure-11} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend.}} \kge{The performance of a set of tuning parameter values often do not agree across metrics. For example, both density methods perform well according to average fidelity but poorly according to MSEE, and all quantile bin methods perform well accroding to average $R^2$ but poorly based on average fidelity and MSEE. However, there is consistency in results across different Gower powers.} The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME. \kge{\sout{Perhaps we can say that the kernel density method or 3 equal bins are the best performing methods, but unfortunately, Figure \ref{fig:figure-12} leaves us without a clear indication of which set of explanations to use to explain the random forest.}} \kge{It is not obvious which set of of tuning parameters are best, but perhaps, we can say that based on all three metrics, the 4 equal bins with a Gower power of 0.5 performs bettter than the other tuning parameter values considered.}

To provide a more detailed view of explanations obtained using different tuning parameter values, we take a closer look at explanations for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match), using explanation scatterplots. Figure~\ref{fig:figure-13} includes the visual representation plots of explanations (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 4 equal bins and 4 quantile bins are included to depict tuning parameter values with good (4 equal bins) and mediocre (4 quantile bins) performance based on the assessment metric plot. Plots for the kernel density simulation method are included in Appendix~\ref{exp-scatter_plus} (Figure \ref{fig:figure-B2}).

First, focus on the explanation for case NM with 4 quantile bins. \kge{The explanation plot from \emph{lime} shows that observed values with rough correlation (rough\_cor) greater than 0.05 and mismatches value less than or equal to 7.17 support a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with rough correlation greater then 0.05 and mismatches less than 7.17 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of cross correlation function (ccf) greater than 0.28 and less than or equal to 0.35 supports a non-match, and the scatter plots shows that the simulated values within this region have mostly random forest predictions less than 0.5.  However, the random forest assigns a probability of 0.19 to case NM, which is not explained by the LIME explanation. Instead, consider that a better explanation based on the scatterplots would be that the random forest assigned a prediction probably close to 0 since the observation has a rough correlation less than 0.5, mismatches greater than 4, and cross correlation function 0.6.} \kge{\sout{Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls in and supports the random forest prediction of a non-match.}}

The other explanation scatterplots can be interpreted in a similar manner. Overall, the plots identify that the \kge{\sout{3} 4} quantile bin explanations are not faithful to the random forest model predictions. The \kge{4 \sout{3}} equal bin explanations are more faithful to the random forest, but the regions captured by the bins do not align \kge{as well as they could} with the regions \kge{containing \sout{with}} similar probabilities produced by the random forest.

@@

\end{knitrout}
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: \kge{4 \sout{3}} quantile bins (first column) and \kge{4 \sout{3}} equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}
